---
title: "R Code for Hoefer et al. 2026 - Passive acoustic monitoring outperforms observer-based methods for Australian frog communities"
author: "Sebastian Hoefer"
date: today
title-block-banner: "#161616"
title-block-banner-color: "white"
format:
  html:
    embed-resources: true
    grid: 
      body-width: 2000px
      sidebar-width: 200px
      margin-width: 300px
    toc: true
    toc-title: \deftocheading{toc}{}
    toc-depth: 6
    toc-location: right
    toc-expand: 1
    code-overflow: wrap
    code-block-bg: "F2F2F2"
    code-block-border-left: "#4A9D9E"
    highlight-style: github
    theme: darkly
    knitr:
      opts_chunk: 
        warning: false
        R.options:
          width: 120
standalone: true
---

# Load packages

Load packages needed for data manipulation and analyses.

```{r}
#| message: false

library(tidyverse)        # Data wrangling and visualization
library(colorspace)       # Color manipulation
library(cmdstanr)         # Interface to 'CmdStan'
library(brms)             # Bayesian regression models
library(rstan)            # Interface to Stan
library(ggeffects)        # Marginal effects plots
library(DHARMa)           # Residual diagnostics
library(emmeans)          # Estimated marginal means
library(tidybayes)        # Tidy data for Bayesian models
library(HDInterval)       # HPD intervals
library(ggsignif)         # Significance levels in ggplot
library(vegan)            # Community ecology analysis
library(pairwiseAdonis)   # Pairwise comparisons using adonis2
library(ggbeeswarm)       # For raw data plotting
library(patchwork)        # Combine ggplots
library(iNEXT)            # Species accumulation curves
library(scales)           # Manipulate axis labels
library(report)           # Automated reporting
```

# Functions

Custom ggplot theme for visualisation.

```{r}
my.theme <- function(){
  theme_classic() +
    theme(text = element_text(family = "Avenir Next"),
          axis.title.y = element_text(margin = margin(t = 0,r = 20,b = 0,l = 0)),
          axis.title.x = element_text(margin = margin(t = 20,r = 0,b = 0,l = 0)), 
          plot.margin = unit(c(5, 10, 5, 10), units = "mm"),
          strip.background = element_rect(fill = "#CCCCFF"),
          strip.text.x = element_text(size = 20),
          axis.title = element_text(size = 20),
          axis.text = element_text(size = 18),
          legend.text = element_text(size = 15),
          legend.title = element_text(size = 15))
}
```

# Load data

## All frog data

```{r}
frog_data <- read.csv("frog_data.csv")
```

Function to summarise the total number of species per method for each plot across all sites. Adding missing zero values for methods that didn't capture any species at a given plot.

```{r}
summarise_frog_data <- function(df) {
  
  site_plots <- c("Tarcutta.DryA", "Tarcutta.DryB", "Tarcutta.WetA", "Tarcutta.WetB", 
                  "Duval.DryA", "Duval.DryB", "Duval.WetA", "Duval.WetB", 
                  "Mourachan.DryA", "Mourachan.DryB", "Mourachan.WetA", "Mourachan.WetB",
                  "Wambiana.DryA", "Wambiana.DryB", "Wambiana.WetA", "Wambiana.WetB",
                  "Undara.DryA", "Undara.DryB", "Undara.WetA", "Undara.WetB",
                  "Rinyirru.DryA", "Rinyirru.DryB", "Rinyirru.WetA", "Rinyirru.WetB")

  sites <- c("Tarcutta", "Duval", "Mourachan", "Wambiana", "Undara", "Rinyirru")
  
  methods <- c("PAM.all", "OBM", "PAM.survey")
  
  df <- df %>%
    unite(site.plot, c(site, plot), sep = ".", remove = FALSE) %>%
    select(-plot) %>%
    group_by(site, site.plot, assessment.method) %>%
    summarise(richness = length(unique(scientific.name))) %>%
    ungroup() %>%
    select(site, site.plot, assessment.method, richness) %>%
    full_join(crossing(site.plot = unique(.$site.plot),
                       assessment.method = unique(.$assessment.method)),
              by = c("site.plot", "assessment.method")) %>%
    replace_na(list(richness = 0)) %>%
    mutate(site.plot = factor(site.plot, levels = site_plots),
           site = str_extract(site.plot, "^[^.]*"),
           site = factor(site, levels = sites),
           assessment.method = factor(assessment.method, levels = methods),
           richness = as.numeric(richness))

  return(df)
}
```

Run function.

```{r}
#| message: false

frog_data_summary <- summarise_frog_data(frog_data)
```

## PAM data

```{r}
pam_data <- read_csv("frogs_PAM_daily_all.csv", show_col_types = FALSE) %>% 
  filter(!MANUAL.ID %in% c("UNSURE", NA)) %>% 
  mutate(MANUAL.ID = ifelse(MANUAL.ID == TRUE, 1, 0)) %>% 
  group_by(template) %>%
  filter(any(MANUAL.ID == 1)) %>% #Filter for templates with at least one true positive
  ungroup()
```

# BirdNET Embeddings Performance Check

## Recall calculations Embeddings

This is a combined approach where if we have enough data for positive and negative detections for a template (>5 detections) we sample from the real distribution of the data (does not assume normality). If we have less than 5 detections for either positive or negative, we don't know the variance, so we make an educated guess based on the other frogs (we use "Parametric Simulation", borrowing the SD from the other templates, to simulate the variation we would expect to see if we had more detections). 

```{r}
#| warning: false

# Function to return predictions over extended range
run_model <- function(df) {
  model_logistic <- glm(MANUAL.ID ~ eucledian.distance, 
                       family = binomial(link = "logit"), 
                       data = df)
  
  # Create extended x range
  new_x <- data.frame(
    eucledian.distance = seq(min(df$eucledian.distance), 
                           max(df$eucledian.distance), 
                           length.out = 100)
  )
  
  # Get predictions and SE
  preds <- predict(model_logistic, newdata = new_x, type = "link", se.fit = TRUE)
  
  # Add predictions to new_x
  new_x$fit <- plogis(preds$fit)
  new_x$lower <- plogis(preds$fit - 1.96 * preds$se.fit)
  new_x$upper <- plogis(preds$fit + 1.96 * preds$se.fit)
  
  return(new_x)  # Return only the predictions dataframe
}

# Get the predictions
model_predictions <- pam_data %>%
  nest(data = -template) %>%               
  mutate(preds = map(data, run_model)) %>%
  select(-data) %>%                       
  unnest(preds) 

# Calculate the average SD from templates that have >1 detection
global_pos_sd <- pam_data %>%
  filter(MANUAL.ID == 1) %>%
  group_by(template) %>%
  summarise(
    sd = sd(eucledian.distance), 
    n = n()
  ) %>%
  filter(n > 1) %>%
  summarise(avg_sd = mean(sd)) %>%
  pull(avg_sd)

calc_prob_hybrid <- function(pos, neg, global_sd, min_n = 5, n_sim = 10000) {
  
  if(length(pos) == 0 || length(neg) == 0) return(NA)
  
  set.seed(123)
  
  # --- HANDLE POSITIVES ---
  if(length(pos) < min_n) {
    # If fewer than 5 detections, use synthetic data.
    # We use the specific mean (best guess for location)
    # But the global SD (best guess for spread)
    pos_draws <- abs(rnorm(n_sim, mean = mean(pos), sd = global_sd))
  } else {
    # STANDARD: If we have enough data, trust the specific data distribution
    pos_draws <- sample(pos, n_sim, replace = TRUE)
  }
  
  # --- HANDLE NEGATIVES ---
  # Assuming negatives usually have plenty of data (background noise)
  if(length(neg) < min_n) {
     neg_global_sd <- sd(neg)
     neg_draws <- abs(rnorm(n_sim, mean = mean(neg), sd = neg_global_sd))
  } else {
     neg_draws <- sample(neg, n_sim, replace = TRUE)
  }
  
  # --- COMPARE ---
  mean(pos_draws < neg_draws)
}

# Apply
template_probabilities <- pam_data %>%
  group_by(template) %>%
  summarise(
    probability = calc_prob_hybrid(
      pos = eucledian.distance[MANUAL.ID == 1],
      neg = eucledian.distance[MANUAL.ID == 0],
      global_sd = global_pos_sd,
      min_n = 5
    )
  )

```

## Visualisation

```{r}
#| fig-width: 27
#| fig-asp: 0.6
#| warning: false

(combined_plot <- ggplot(pam_data, aes(x = eucledian.distance, y = MANUAL.ID)) +  
  # Add CI ribbon
  geom_ribbon(data = model_predictions,
              aes(y = fit, ymin = lower, ymax = upper),
              fill = "grey70",
              alpha = 0.3) +
  # Add CI boundary lines
  geom_line(data = model_predictions,
            aes(y = lower),
            linetype = "dashed",
            color = "black",
            linewidth = 0.8) +
  geom_line(data = model_predictions,
            aes(y = upper),
            linetype = "dashed",
            color = "black",
            linewidth = 0.8) +
  # Add mean prediction line
  geom_line(data = model_predictions,
            aes(y = fit),
            color = "black",
            linewidth = 2) +
  geom_rug(data = subset(pam_data, MANUAL.ID == 0),
           aes(x = eucledian.distance),
           color = "darkorchid", alpha = 0.3,
           length = unit(0.05, "npc"),
           linewidth = 0.8, 
           sides = "b",
           inherit.aes = FALSE) +
  geom_rug(data = subset(pam_data, MANUAL.ID == 1),
           aes(x = eucledian.distance),
           color = "darkseagreen", alpha = 0.3,
           length = unit(0.05, "npc"),
           linewidth = 0.8, 
           sides = "t",
           inherit.aes = FALSE) +
  geom_text(data = template_probabilities,
            aes(x = Inf, y = Inf, label = sprintf("%.2f", probability)),
            hjust = 1.3, vjust = 2, inherit.aes = FALSE) +
  facet_wrap(~ template, scales = "free_x", ncol = 10) +
  labs(x = "Euclidean Distance", 
       y = "Probability of True Positive") +
  scale_y_continuous(limits = c(-0.1, 1.1), breaks = seq(0, 1, 0.25)) +
  theme_bw() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.title = element_text(size = 23, face = "bold"),
        axis.title.x = element_text(margin = margin(t = 25, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 25, b = 0, l = 0)),
        plot.margin = unit(c(5, 10, 5, 10), units = "mm")))
```

# OBM individually

```{r}
summarise_frog_data_methods <- function(df) {
  
  site_plots <- c("Tarcutta.DryA", "Tarcutta.DryB", "Tarcutta.WetA", "Tarcutta.WetB", 
                  "Duval.DryA", "Duval.DryB", "Duval.WetA", "Duval.WetB", 
                  "Mourachan.DryA", "Mourachan.DryB", "Mourachan.WetA", "Mourachan.WetB",
                  "Wambiana.DryA", "Wambiana.DryB", "Wambiana.WetA", "Wambiana.WetB",
                  "Undara.DryA", "Undara.DryB", "Undara.WetA", "Undara.WetB",
                  "Rinyirru.DryA", "Rinyirru.DryB", "Rinyirru.WetA", "Rinyirru.WetB")

  sites <- c("Tarcutta", "Duval", "Mourachan", "Wambiana", "Undara", "Rinyirru")
  
  methods <- c("funnel", "pitfall","spotlighting","incidentals","cover board","PAM.survey","PAM.all")
  
  df <- df %>%
    unite(site.plot, c(site, plot), sep = ".", remove = FALSE) %>%
    select(-plot) %>%
    group_by(site, site.plot, survey.method) %>%
    summarise(richness = length(unique(scientific.name))) %>%
    ungroup() %>%
    select(site, site.plot, survey.method, richness) %>%
    full_join(crossing(site.plot = unique(.$site.plot),
                       survey.method = unique(.$survey.method)),
              by = c("site.plot", "survey.method")) %>%
    replace_na(list(richness = 0)) %>%
    mutate(site.plot = factor(site.plot, levels = site_plots),
           site = str_extract(site.plot, "^[^.]*"),
           site = factor(site, levels = sites),
           survey.method = factor(survey.method, levels = methods),
           richness = as.numeric(richness))

  return(df)
}
```

Run function.

```{r}
#| message: false

frog_data_summary_methods <- summarise_frog_data_methods(frog_data)
```

```{r}
#| fig-width: 12
#| fig-asp: 0.8
#| warning: false

pal <- c("#D03C00","#F78303","black", "#0CC170", "#7671B4","#189F9F","#18709F")
pal_dark <- after_scale(darken(pal, .1, space = "HLS"))

(survey.methods.pointrange <- ggplot(frog_data_summary_methods, aes(survey.method, richness)) +
   
   # 1. Raw Data (Background)
   # We keep the raw points faint so the viewer can still see the sample size/spread
   geom_point(aes(fill = survey.method), 
              color = "transparent",
              position = position_jitter(width = 0.2, height = 0, seed = 123),
              shape = 21, stroke = 0, size = 2.5, alpha = 0.25) +
   
   # 2. The Pointrange (Mean +/- Standard Error)
   # We use stat_summary so we don't have to pre-calculate the means
   stat_summary(
     aes(color = survey.method, fill = after_scale(desaturate(lighten(color, .6), .4))),
     fun.data = mean_cl_boot,       # Calculates 95% CI
     geom = "pointrange",
     size = 1,
     linewidth = 1.2,
     shape = 21,
     stroke = 1.5
   ) +
   my.theme() +
   scale_y_continuous(limits = c(0, 14), breaks = seq(0, 14, by = 2),
                      name = "Total Species Richness") +
   scale_x_discrete(name = "", labels = c("Funnel trap", "Pitfall trap", "Spotlighting", 
                                          "Incidental","Arboreal Cover Board" ,"PAM (survey period)", "PAM (all audio)")) +
   scale_fill_manual(values = pal, guide = "none") +
   scale_colour_manual(values = pal_dark, name = "") +
   theme(legend.position = "none",
         axis.text.x = element_text(angle = 45, hjust = 1)))

```

# Financial investment

```{r}
costs_daily <- read_csv("expenses_daily.csv") %>% 
  mutate(date = dmy(date)) %>% 
  filter(method != "camera") %>% 
  group_by(method) %>%
  arrange(date) %>%
  mutate(cumulative_cost = cumsum(cost),
         method = factor(method, levels = c("OBM", "PAM"))) %>% 
  ungroup() %>% 
  drop_na(method)
```
```{r}
#| fig-width: 14
#| fig-asp: 0.6
#| warning: false

(costs <- ggplot(costs_daily, aes(date, cumulative_cost, col = method, group = method)) +
  annotate("rect", xmin = as.Date("2021-04-16"), xmax = as.Date("2021-07-13"), ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.3) +
  annotate("rect", xmin = as.Date("2021-09-26"), xmax = as.Date("2021-11-16"), ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.3) +
  annotate("rect", xmin = as.Date("2022-04-26"), xmax = as.Date("2022-08-15"), ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.3) +
  annotate("rect", xmin = as.Date("2022-09-26"), xmax = as.Date("2022-11-30"), ymin = -Inf, ymax = Inf, fill = "grey", alpha = 0.3) +
  geom_line(size = 2) +
  my.theme() +
  scale_y_continuous(name = "Cost in AUD",
                     limits = c(0, 245000),
                     breaks = seq(0, 250000, by = 50000),
                     labels = label_dollar(prefix = "$", big.mark = ",")) +
  scale_x_date(name = "", date_breaks = "1 month", expand = c(0.002,0.02), date_labels = "%b %Y") +
  scale_colour_manual(values=c("#FF8C00", "#189F9F"),
                      name = "",
                      label = c("OBM", "PAM")) +
  theme(legend.position = c(y=0.05, x=0.9),
        legend.background = element_blank(),
        legend.key = element_blank(),
        panel.grid.major.y = element_line(colour = "black", size = 0.4, linetype = "dotted"),
        axis.text.x = element_text(angle = 45, hjust = 1)))
```

# Species Richness

## Bayesian Analysis

### Priors

Defining priors for Bayesian models.

::: {.panel-tabset .nav-pills}

#### Priors for the intercept

```{r}
frog_data_summary %>% summarise(median(log(richness)))
frog_data_summary %>% summarise(mad(log(richness)))
```

Priors for the intercept are 1.6 and 1.

#### Priors for the slope

```{r}
log(sd(frog_data_summary$richness))/apply(model.matrix(~assessment.method*site, data = frog_data_summary), 2, sd)
```

Priors for the slope are 5.
:::

### Fit model

::: {.panel-tabset .nav-pills}

#### Model 1

```{r}
priors1 <- prior(normal(1.6,1), class  = "Intercept") +
  prior(normal(0,5), class = "b") +
  prior(student_t(3,0,5), class = "sd")

methods.form1 <- bf(richness ~ assessment.method + (1|site/site.plot),
                 family = poisson(link = "log"))

methods.brm1 <- brm(methods.form1,
                  data = frog_data_summary,
                  prior = priors1,
                  sample_prior = "yes",
                  backend = 'cmdstanr',
                  threads = threading(2),
                  chains = 4,
                  cores = 4,
                  iter = 2000,
                  warmup = 1000,
                  thin = 1,
                  refresh = 0,
                  seed = 3,
                  save_pars = save_pars(all = TRUE),
                  control = list(adapt_delta = 0.90, max_treedepth = 12))
```

##### Predictions

```{r}
methods.brm1 %>% ggpredict() %>% plot()
```

#### Model 2

```{r}
priors2 <- prior(normal(1.6,1), class  = "Intercept") +
  prior(normal(0,5), class = "b") +
  prior(student_t(3,0,5), class = "sd")

methods.form2 <- bf(richness ~ assessment.method + (1|site/site.plot),
                 family="negbinomial2")

methods.brm2 <- brm(methods.form2,
                  data = frog_data_summary,
                  prior = priors2,
                  sample_prior = "yes",
                  backend = 'cmdstanr',
                  threads = threading(2),
                  chains = 4,
                  cores = 4,
                  iter = 2000,
                  warmup = 1000,
                  thin = 1,
                  refresh = 0,
                  seed = 3,
                  save_pars = save_pars(all = TRUE),
                  control = list(adapt_delta = 0.95, max_treedepth = 12))
```

##### Predictions

```{r}
methods.brm2 %>% ggpredict() %>% plot()
```

:::

### Compare Models

::: {.panel-tabset .nav-pills}

##### Loo Model 1

```{r}
(l.1 <- methods.brm1 %>% loo())
```

##### Loo Model 2
```{r}
(l.2 <- methods.brm2 %>% loo())
```

##### Loo Comparison
```{r}
loo_compare(loo(methods.brm1), loo(methods.brm2))
```

Model `methods.brm1` was selected as best model based on loo estimates.

:::

### Diagnostics

::: {.panel-tabset .nav-pills}

##### Trace plots

```{r}
methods.brm1$fit %>% stan_trace()
```

##### Autocorrelation plots

```{r}
methods.brm1$fit %>% stan_ac()
```

##### Rhat statistic

```{r}
methods.brm1$fit %>% stan_rhat()
```

##### Effective sampling size

```{r}
methods.brm1$fit %>% stan_ess()
```

##### Posterior predictive check plot

```{r}
methods.brm1 %>% pp_check(type = "dens_overlay", ndraws = 200)
```

##### DHARMa residuals

```{r}
set.seed(8)
preds <- posterior_predict(methods.brm1, ndraws=250, summary=FALSE)
method.resids <- createDHARMa(simulatedResponse = t(preds),
                            observedResponse = frog_data_summary$richness,
                            fittedPredictedResponse = apply(preds, 2, median),
                            integerResponse = TRUE)
method.resids %>% plot()
```

##### Dispersion test

```{r}
method.resids %>% testDispersion()
```

##### Zero-inflation test

```{r}
method.resids %>% testZeroInflation()
```

:::

### Investigation

```{r}
summary(methods.brm1)
```

#### Methods pairwise comparison across sites

```{r}
(diff.methods.avg <- methods.brm1 %>%
  emmeans(~assessment.method) %>%
  regrid(transform = "none") %>%
  pairs() %>% 
  gather_emmeans_draws() %>% 
  mutate(f.change = exp(.value)) %>% 
  summarise("Average fractional change" = median(f.change),
            "Lower HDI" = HDInterval::hdi(f.change)[1],
            "Upper HDI" = HDInterval::hdi(f.change)[2],
            "Probability of difference" = sum(.value > 0)/n())) #to see if there is any change
```

### Visualisation

```{r}
#| fig-width: 12
#| fig-asp: 0.6

pal <- c("#189F9F","#FF8C00", "#18709F", "#A034F0")
pal_dark <- after_scale(darken(pal, .1, space = "HLS"))

(methods.pointrange <- ggplot(frog_data_summary, aes(assessment.method, richness)) +
   geom_point(aes(fill = assessment.method), 
              color = "transparent",
              position = position_jitter(width = 0.2, height = 0, seed = 123),
              shape = 21, stroke = 0, size = 3, alpha = 0.25) +
   stat_summary(
     aes(color = assessment.method, fill = after_scale(desaturate(lighten(color, .6), .4))),
     fun.data = mean_cl_boot,
     geom = "pointrange",
     size = 1.2,             
     linewidth = 1.2,        
     shape = 21,             
     stroke = 1.5            
   ) +
   geom_signif(comparisons = list(c("PAM.all", "OBM"),
                                  c("PAM.all", "PAM.survey"),
                                  c("OBM", "PAM.survey")),
               annotations = c("1.65, HDI 95% [1.33,2.05]",
                               "2.8, HDI 95% [2.12,3.58]",
                               "1.7, HDI 95% [1.26,2.20]"),
               y_position = c(14, 16, 11),
               textsize = 6, vjust = -0.2) +
   my.theme() +
   scale_y_continuous(limits = c(0, 18), breaks = seq(0, 18, by = 2),
                      name = "Total Species Richness") +
   scale_x_discrete(name = "", 
                    labels = c("PAM (all audio)", "OBM", "PAM (survey period)")) +
   scale_fill_manual(values = pal, guide = "none") +
   scale_colour_manual(values = pal_dark, guide = "none") +
   theme(legend.position = "none"))
```

# Community Composition

## Prepare data

```{r}
frog_community3 <- frog_data %>% 
  filter(!recapture %in% c("y")) %>%  
  select(site, plot, assessment.method,scientific.name) %>% #selected relevant grouping variables
  group_by(across(everything())) %>% #groups by all available columns
  mutate(number = n()) %>% #created a numbers column that counts the number of duplicate observations
  ungroup() %>% 
  unique() %>% #merged the duplicate rows
  pivot_wider(names_from = "scientific.name", #transformed dataset from long to wide format
              values_from = "number",
              values_fill = list(number=0)) %>% #replaced NAs with 0
  arrange(site) %>% 
  mutate(assessment.method = factor(assessment.method))
```

In preparation for community analyses, only keep columns and rows with values above 0.

```{r}
all <- frog_community3 %>% select(where(~ any(. != 0))) %>% filter(rowSums(.[,4:ncol(.)]) > 0)
```

Transformation of the data to binary (presence/absence) for Jaccard dissimilarity.

```{r}
all_jacc <- all %>%
  select(where(is.numeric)) %>% # Removed columns with non-numerical values
  decostand(method="pa", MARGIN = 1) %>% # Standardised presence/absence (method = "pa") by rows (MARGIN = 1)
  cbind(all[,1:3]) %>% 
  select(site, plot, assessment.method, everything())
```

## NMDS Ordination

### Jaccard dissimilarity

```{r}
set.seed(3)
all_nmds_jacc <-  metaMDS(all_jacc[4:ncol(all_jacc)], distance="jaccard", k=2,trymax=1000, autotransform = FALSE)
```

## Visualisation

### Jaccard dissimilarity

```{r}
#| fig-width: 14
#| fig-asp: 0.6

# 1. Extract NMDS Scores from your specific object (all_nmds_jacc)
# We pull the metadata columns (site, plot, assessment.method) from 'all_jacc'
scores_df <- as.data.frame(scores(all_nmds_jacc, display = "sites")) %>%
  bind_cols(all_jacc[, 1:3]) %>%  # Bind the first 3 cols: site, plot, method
  mutate(assessment.method = factor(assessment.method, 
                                    levels = c("PAM.all", "OBM", "PAM.survey")),
         site = factor(site, 
                                    levels = c("Rinyirru", "Undara", "Wambiana", "Mourachan", "Duval", "Tarcutta")))

# 2. Calculate Centroids (Means for each method)
centroids <- scores_df %>%
  group_by(assessment.method) %>%
  summarise(NMDS1 = mean(NMDS1), NMDS2 = mean(NMDS2), .groups = "drop")

# 3. Create the Plot
pal <- c("#189F9F","#FF8C00", "#18709F")
pal_dark_values <- darken(pal, 0.1, space = "HLS")
pal_light_fill <- desaturate(lighten(pal_dark_values, .6), .4)

(gg_nmds_ellipses_plot <- ggplot(scores_df, aes(x = NMDS1, y = NMDS2, color = assessment.method)) +
   stat_ellipse(aes(fill = assessment.method), 
                geom = "polygon", 
                level = 0.95, 
                alpha = 0.1,
                linetype = "dashed",
                show.legend = FALSE) +
   geom_point(size = 2, alpha = 0.4, shape = 16, show.legend = FALSE) +
   geom_point(data = centroids,
              aes(color = assessment.method,
                  fill = after_scale(desaturate(lighten(color, .6), .4))),
              size = 6, shape = 21, stroke = 1) +
    annotate("text", x = Inf, y = Inf, 
            label = paste("Stress:", round(all_nmds_jacc$stress, 3)), 
            hjust = 1.2, vjust = 2, size = 5, fontface = "italic") +
   scale_color_manual(values = pal_dark_values,
                      name = "",
                      labels = c("PAM (all audio)", "OBM", "PAM (survey period)")) +
   scale_fill_manual(values = pal,
                     guide = "none") +
    guides(color = guide_legend(override.aes = list(
     shape = 21,              # Force Circle
     size = 6,                # Match Centroid Size
     fill = pal_light_fill,   # Use the PRE-CALCULATED light colors
     color = pal_dark_values, # Use the dark borders
     stroke = 1,
     alpha = 1                # Make sure it's visible (override the alpha=0 from dummy layer)
   ))) +
   my.theme() +
   theme(
     panel.grid = element_blank(),
     legend.position = "bottom",
     legend.text = element_text(size = 16),
     axis.title = element_text(face = "bold", size = 12)
   ))
```

### Investigation

#### Adonis

```{r}
#| warning: false

#Permutation test: significant difference between assessment methods for frog communities.
all.dist2 <- vegdist(all_jacc[4:ncol(all_jacc)], 'jaccard')
(all.adonis <- adonis2(all.dist2 ~ assessment.method, data=all_jacc, strata = all_jacc$site))

#Let's see how each method compares within site
set.seed(9)
raw_results <- pairwise.adonis2(all.dist2 ~ assessment.method, 
                                data = all_jacc,
                                strata = "site")

(results_df <- data.frame(
  pair = names(raw_results)[-1],
  F_stat = sapply(raw_results[-1], function(x) x$F[1]),
  R2     = sapply(raw_results[-1], function(x) x$R2[1]),
  p_raw  = sapply(raw_results[-1], function(x) x$`Pr(>F)`[1])
) %>%
  mutate(p_adjusted = p.adjust(p_raw, method = "holm")))

```

#### Magnitude of differences

```{r}
# Create Community Matrix (Wide Format)
frog_comm_plot <- frog_data %>% 
  filter(!recapture %in% c("y")) %>% 
  select(site, plot, assessment.method, scientific.name) %>%
  group_by(site, plot, assessment.method, scientific.name) %>%
  summarise(present = 1, .groups = "drop") %>% 
  pivot_wider(names_from = scientific.name,
              values_from = present,
              values_fill = 0) %>%
  select(where(~any(. != 0))) %>% 
  filter(rowSums(select(., where(is.numeric))) > 0)

# Separate Metadata for joining later
frog_meta <- frog_comm_plot %>%
  select(site, plot, assessment.method) %>%
  mutate(row_id = as.character(row_number())) # ID for re-joining

```

```{r}
# Calculate Distance Matrix
dist_matrix <- frog_comm_plot %>%
  select(where(is.numeric)) %>%
  vegdist(method = "jaccard")

# Convert to Long Format
dist_long <- dist_matrix %>%
  as.matrix() %>%
  as_tibble(rownames = "sample1") %>%
  pivot_longer(-sample1, names_to = "sample2", values_to = "distance") %>%
  filter(sample1 < sample2) # Keep upper triangle only (no duplicates)

# Join Metadata & Define Comparison Types
dist_categorized <- dist_long %>%
  # Join metadata for Sample 1
  left_join(frog_meta, by = c("sample1" = "row_id")) %>%
  # Join metadata for Sample 2 (with suffixes)
  left_join(frog_meta, by = c("sample2" = "row_id"), 
            suffix = c("_1", "_2")) %>%
  
  # Logic to define what we are comparing
  mutate(
    comparison_type = case_when(
      # 1. METHOD EFFECT: Same Site, Same Plot, Different Method
      # (How much does changing the method change the result?)
      site_1 == site_2 & plot_1 == plot_2 & assessment.method_1 != assessment.method_2 ~ "Method Effect",
      
      # 2. PLOT TURNOVER: Same Site, Different Plot, Same Method
      # (How much does moving 500m change the result?)
      site_1 == site_2 & plot_1 != plot_2 & assessment.method_1 == assessment.method_2 ~ "Plot Turnover\n(Within Site)",
      
      # 3. SITE TURNOVER: Different Site, Same Method
      # (How much does moving 100km change the result?)
      site_1 != site_2 & assessment.method_1 == assessment.method_2 ~ "Site Turnover\n(Between Sites)"),
    site_1 = fct_relevel(site_1, "Rinyirru", "Undara", "Wambiana", "Mourachan", "Duval", "Tarcutta")) %>%
  # Remove comparisons we don't care about (e.g. Diff Site AND Diff Method)
  drop_na(comparison_type)

```

##### Visualisation

```{r}
#| fig-width: 10
#| fig-asp: 0.6

(jacc_plot_overall <- ggplot(dist_categorized, aes(x = comparison_type, y = distance)) +
   ggbeeswarm::geom_quasirandom(aes(color = comparison_type),
                    size = 1.3,
                    width = .2,
                    alpha = .3,
                    show.legend = FALSE) +
   stat_summary(fun.data = mean_cl_normal,
                geom = "pointrange",
                color = "black",
                size = 1.1,
                linewidth = 1.1) +
   scale_color_manual(values = c("grey40", "grey40", "grey40")) + # Highlight Method Effect in Red
   scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
   my.theme() +
   theme(
     panel.border = element_rect(fill = NA, color = "black", linewidth = 1.5)
   ) +
   labs(x = "",
        y = "Jaccard Dissimilarity"))

```

```{r}
#| fig-width: 10
#| fig-asp: 0.6

(jacc_plot_methods_by_site <- dist_categorized %>%
   filter(comparison_type == "Method Effect") %>%
   ggplot(aes(x = site_1, y = distance)) +
   ggbeeswarm::geom_quasirandom(color = "grey40",
                    size = 1.3,
                    width = .2,
                    alpha = .5) +
   stat_summary(fun.data = mean_cl_normal,
                geom = "pointrange",
                color = "black",
                size = 1.1,
                linewidth = 1.1) +
   scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
   my.theme() +
   theme(
     panel.border = element_rect(fill = NA, color = "black", linewidth = 1.5),
     axis.text.x = element_text(angle = 45, hjust = 1)
   ) +
   labs(x = "",
        y = "Jaccard Dissimilarity"))
```

```{r}
#| fig-width: 14
#| fig-asp: 1.2

gg_nmds_ellipses_plot / jacc_plot_overall + jacc_plot_methods_by_site +
  plot_annotation(tag_levels = "A", tag_suffix = ")") &
  theme(plot.tag = element_text(size = 35, face = "bold"),
        plot.margin = unit(c(5, 10, 0, 15), units = "mm"),
        plot.tag.position = c(-0.05, 1))
```

Split the NMDS plot into individual sites.

```{r}
#| fig-width: 14
#| fig-asp: 0.6

# 2. Calculate Centroids PER SITE and PER METHOD
site_method_centroids <- scores_df %>%
  group_by(site, assessment.method) %>%
  summarise(NMDS1 = mean(NMDS1), NMDS2 = mean(NMDS2), .groups = "drop")

# Calculate mean Method Effect dissimilarity per site
site_jacc_labels <- dist_categorized %>%
  filter(comparison_type == "Method Effect") %>%
  group_by(site = site_1) %>%
  summarise(
    mean_dist = round(mean(distance), 2),
    .groups = "drop"
  ) %>%
  mutate(
    label_text = paste0("Mean Dissim: ", mean_dist)
  )


# 3. Create the Faceted Plot
(gg_nmds_ellipses_plot_sites <- ggplot(scores_df, aes(x = NMDS1, y = NMDS2, color = assessment.method)) +
   stat_ellipse(aes(fill = assessment.method), 
                geom = "polygon", 
                level = 0.95, 
                alpha = 0.1,
                linetype = "dashed",
                show.legend = FALSE) +
   geom_point(size = 2, alpha = 0.4, shape = 16, show.legend = FALSE) +
   geom_point(data = site_method_centroids,
              aes(color = assessment.method,
                  fill = after_scale(desaturate(lighten(color, .6), .4))),
              size = 6, shape = 21, stroke = 1) +
   facet_wrap(~site, scales = "free") +
   geom_text(data = site_jacc_labels,
             aes(label = label_text),
             x = Inf, y = Inf,
             hjust = 1.1, vjust = 1.5,
             color = "black",    
             size = 3.5, fontface = "italic",
             inherit.aes = FALSE) +
   scale_color_manual(values = pal_dark_values,
                      name = "",
                      labels = c("PAM (all audio)", "OBM", "PAM (survey period)")) +
   scale_fill_manual(values = pal,
                     guide = "none") +
    labs(subtitle = paste("Global NMDS Stress:", round(all_nmds_jacc$stress, 3))) +
    guides(color = guide_legend(override.aes = list(
     shape = 21,              # Force Circle
     size = 6,                # Match Centroid Size
     fill = pal_light_fill,   # Use the PRE-CALCULATED light colors
     color = pal_dark_values, # Use the dark borders
     stroke = 1,
     alpha = 1                # Make sure it's visible (override the alpha=0 from dummy layer)
   ))) +
   my.theme() +
   theme(
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1.5),
    strip.text.x = element_text(size = 14, face = "bold", hjust = 0),
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(face = "italic", size = 13)))
```

# Survey Effort

## Species accumulation curves

### Prepare data

Merged duplicate rows for duplicate observations and transformed the dataset to a wide format.

```{r}
frog_community <- frog_data %>% 
  select(site, SamplingDay, assessment.method,scientific.name) %>% #selected relevant grouping variables
  group_by(across(everything())) %>% #groups by all available columns
  mutate(number = n()) %>% #created a numbers column that counts the number of duplicate observations
  ungroup() %>% 
  unique() %>% #merged the duplicate rows
  pivot_wider(names_from = "scientific.name", #transformed dataset from long to wide format
              values_from = "number",
              values_fill = list(number=0)) %>% #replaced NAs with 0
  arrange(site) %>% 
  filter(!assessment.method %in% c("PAM.all", "camera")) %>% 
  mutate(SamplingDay = as.factor(SamplingDay))

frog_community2 <- frog_community %>%
    group_by(site, SamplingDay) %>%
    summarise(
        assessment.method = "total",
        across(where(is.numeric), sum),
        .groups = "drop"
    ) %>%
    bind_rows(frog_community, .)
```

```{r}
# Prepare Incidence Data (Days as Effort)
incidence_data <- frog_community2 %>%
  unite("Group_ID", site, assessment.method, sep = "__", remove = FALSE) %>%
  group_by(Group_ID, site, assessment.method) %>%
  mutate(across(where(is.numeric) & !c(SamplingDay), ~ ifelse(. > 0, 1, 0))) %>%
  summarise(
    T_days = case_when(
      unique(site) %in% c("Duval", "Mourachan") ~ 21,
      unique(site) %in% c("Rinyirru", "Undara", "Wambiana", "Tarcutta") ~ 28,
      TRUE ~ as.numeric(n_distinct(SamplingDay))
    ),
    across(where(is.numeric) & !c(SamplingDay), sum),
    .groups = "drop"
  )

# Create iNEXT Input List
list_data <- split(incidence_data, incidence_data$Group_ID)

clean_vector <- function(x) {
  # Extract numeric columns
  vec <- x %>% select(where(is.numeric)) %>% unlist() %>% as.numeric()
  
  # vec[1] is T_days. The rest are species counts.
  t_val <- vec[1]
  species_freqs <- vec[-1]
  
  # Return: c(T_days, count1, count2...) keeping only non-zero species
  return(c(t_val, species_freqs[species_freqs > 0]))
}

iNEXT_input <- lapply(list_data, clean_vector)
```

### Run Analysis

```{r}
#| warning: false

# q=0 is Richness. incidence_freq ensures X-axis is Days.
# endpoint = 90 ensures the extrapolation lines all go out to 90 days
out <- iNEXT(iNEXT_input, q = 0, datatype = "incidence_freq", endpoint = 60, knots = 60)

# Extract Data Frame for Custom Plotting
ggi <- out$iNextEst$size_based %>%
  as_tibble() %>%
  filter(Order.q == 0) %>%
  rename(x = t, y = qD, y.lwr = qD.LCL, y.upr = qD.UCL) %>%
  separate(Assemblage, into = c("site", "assessment.method"), sep = "__") %>%
  mutate(
    site = factor(site, levels = c("Rinyirru","Undara","Wambiana","Mourachan","Duval","Tarcutta")),
    assessment.method = factor(assessment.method, levels = c("PAM.survey", "OBM", "total")),
    Method = ifelse(Method == "Observed", "Rarefaction", Method),
    Method = factor(Method, levels = c("Rarefaction", "Extrapolation"))
  )

# Separate "Total" from "Other Methods"
df_total <- ggi %>% filter(assessment.method == "total")
df_others <- ggi %>% filter(assessment.method != "total")

# Apply Constraint using Interpolation
# We calculate the specific 'Total' limit for every 'x' value in the sub-methods
df_others_constrained <- df_others %>%
  group_by(site) %>%
  mutate(
    # Interpolate the 'y' (Richness) limit
    limit_y = approx(x = df_total$x[df_total$site == unique(site)], 
                     y = df_total$y[df_total$site == unique(site)], 
                     xout = x, rule = 2)$y,
    
    # Interpolate the 'y.upr' (Confidence Interval) limit
    limit_upr = approx(x = df_total$x[df_total$site == unique(site)], 
                       y = df_total$y.upr[df_total$site == unique(site)], 
                       xout = x, rule = 2)$y,
    
    # Clamp the values: Take the minimum of the Method vs the Interpolated Total
    y = pmin(y, limit_y, na.rm = TRUE),
    y.upr = pmin(y.upr, limit_upr, na.rm = TRUE)
  ) %>%
  select(-limit_y, -limit_upr) # Clean up helper columns

# Recombine
ggi_final <- bind_rows(df_total, df_others_constrained) 
```

### Visualisation

```{r}
#| fig-width: 14
#| fig-asp: 0.6


(rare.inext <- ggplot(ggi, aes(x = x, y = y, colour = assessment.method)) +
  geom_ribbon(aes(ymin = y.lwr, ymax = y.upr, fill = assessment.method),
              alpha = 0.15, colour = NA) +
  geom_line(aes(linetype = Method), linewidth = 1) +
  geom_vline(xintercept = c(7, 14, 21, 28), linetype = "dotted", 
             colour = "black", linewidth = 0.3) +
  facet_wrap(~ site, scales = "free_y") +
  scale_x_continuous(name = "Survey effort (days)", 
                     breaks = seq(0, 60, 7), 
                     limits = c(0, 60)) +
  scale_y_continuous(name = "Species Richness", breaks = scales::breaks_pretty(n=4)) +
  scale_colour_manual(
    name = "",
    values = c("PAM.survey" = "#18709F",
               "OBM"        = "#FF8C00",
               "total"      = "black"),
    labels = c("PAM (survey period)",
               "OBM",
               "Total Richness")
  ) +
  scale_fill_manual(
    guide = "none",
    values = c("PAM.survey" = "#18709F",
               "OBM"        = "#FF8C00",
               "total"      = "black") 
  ) +
  scale_linetype_manual(values = c(Rarefaction = "solid", Extrapolation = "22"),
                        guide = guide_legend(title = NULL)) +
  my.theme() +
  theme(
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1.5),
    strip.text.x = element_text(size = 15, face = "bold", hjust = 0),
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(face = "italic", size = 14)
  ) +
  guides(colour = guide_legend(order = 1),
         linetype = guide_legend(order = 2)))
```

## PAM survey length

How many days until PAM.all reaches values similar to OBM for frog species richness. In other words how long should we leave recorders out on average until we get the most adequate number of species.

```{r}
all_frogs <- frog_data %>% 
  filter(assessment.method == "PAM.all") %>% 
  select(site, SamplingDay, assessment.method,scientific.name) %>% #selected relevant grouping variables
  group_by(across(everything())) %>% #groups by all available columns
  mutate(number = n()) %>% #created a numbers column that counts the number of duplicate observations
  ungroup() %>% 
  unique() %>% #merged the duplicate rows
  pivot_wider(names_from = "scientific.name", #transformed dataset from long to wide format
              values_from = "number",
              values_fill = list(number=0)) %>% #replaced NAs with 0
  arrange(site) %>%
  mutate(SamplingDay = as.factor(SamplingDay))
```

```{r}
#| warning: false

# Prepare Incidence Data
incidence_data_all <- all_frogs %>%
  group_by(site) %>%
  mutate(across(where(is.numeric) & !c(SamplingDay), ~ ifelse(. > 0, 1, 0))) %>%
  summarise(
    T_days = case_when(
      unique(site) == "Tarcutta" ~ 986,
      unique(site) == "Duval" ~ 679,
      unique(site) == "Mourachan" ~ 783,
      unique(site) == "Wambiana" ~ 1235,
      unique(site) == "Undara" ~ 780,
      unique(site) == "Rinyirru" ~ 488,
      TRUE ~ n_distinct(SamplingDay)
    ),
    across(where(is.numeric) & !c(SamplingDay), sum),
    .groups = "drop"
  )

# Create Clean Vectors ---
list_data <- split(incidence_data_all, incidence_data_all$site)

clean_vector <- function(x) {
  vec <- x %>% select(where(is.numeric)) %>% unlist() %>% as.numeric()
  return(c(vec[1], vec[-1][vec[-1] > 0])) # T_days, followed by species counts
}

iNEXT_input_list <- lapply(list_data, clean_vector)

# Run iNEXT for Each Site Individually

# Define Max Days for lookup
site_vals <- c(Tarcutta=986, Duval=679, Mourachan=783, Wambiana=1235, Undara=780, Rinyirru=488)

# Function to run iNEXT for one site
run_inext_single <- function(site_name, data_vec) {
  
  # Get the max days for this specific site
  max_d <- site_vals[[site_name]]
  
  # Create the specific sequence of days (1, 2, 3 ... Max)
  day_seq <- seq(1, max_d, by = 1)
  
  # Run iNEXT. Note: We pass the data as a list to keep format consistent
  # size = day_seq forces calculation for every single day
  res <- iNEXT(list(data_vec), q = 0, datatype = "incidence_freq", size = day_seq)
  
  # Extract the dataframe immediately
  df <- res$iNextEst$size_based %>%
    as_tibble() %>%
    filter(Order.q == 0) %>%
    mutate(site = site_name) # Add site name back in
    
  return(df)
}

# Apply the function to all sites and combine results
# map2_dfr iterates over the Names and the Data simultaneously
all_frogs.result <- map2_dfr(names(iNEXT_input_list), iNEXT_input_list, run_inext_single) %>%
  rename(day = t, richness = qD, lower.ci = qD.LCL, upper.ci = qD.UCL) %>%
  mutate(site = factor(site, levels = c("Rinyirru", "Undara", "Wambiana", "Mourachan", "Duval", "Tarcutta")))

obm.max <- ggi_final %>% 
  filter(assessment.method == "OBM",
         Method == "Rarefaction") %>% 
  group_by(site) %>% 
  slice_max(order_by = x, n = 1) %>% 
  ungroup() %>% 
  select(site, richness.max = y)

# Find Days to Reach OBM
days_to_reach <- all_frogs.result %>%
  left_join(obm.max, by = "site") %>%
  group_by(site) %>%
  filter(round(richness, digits = 2) >= richness.max) %>%
  slice(1) %>%
  select(site, days_required = day, pam_richness = richness, obm_target = richness.max)
```

```{r}
pam.max <- all_frogs.result %>% 
  group_by(site) %>% 
  slice_max(order_by = day, n = 1) %>% 
  ungroup() %>% 
  select(site, richness.max = richness)

days.pam.max <- all_frogs.result %>% 
  mutate(richness = round(richness, digits = 1)) %>% 
  group_by(site) %>% 
  slice_max(order_by = richness, n = 1, with_ties = FALSE) %>% 
  ungroup() %>% 
  select(site, richness.max = richness, day)

pam.28 <- all_frogs.result %>% 
  filter(day == 28) %>% 
  select(site, richness.28 = richness) %>% 
  left_join(pam.max, by = "site") %>% 
  mutate(prop.28 = round(richness.28 / richness.max, digits = 2))

pam.90 <- all_frogs.result %>% 
  filter(day == 90) %>% 
  select(site, richness.90 = richness) %>% 
  left_join(pam.max, by = "site") %>% 
  mutate(prop.90 = round(richness.90 / richness.max, digits = 2))

pam.180 <- all_frogs.result %>% 
  filter(day == 180) %>% 
  select(site, richness.180 = richness) %>% 
  left_join(pam.max, by = "site") %>% 
  mutate(prop.180 = round(richness.180 / richness.max, digits = 2))

pam.365 <- all_frogs.result %>% 
  filter(day == 365) %>% 
  select(site, richness.365 = richness) %>% 
  left_join(pam.max, by = "site") %>% 
  mutate(prop.365 = round(richness.365 / richness.max, digits = 2))

pam.rich.prop <- list(pam.max, pam.28, pam.90, pam.180, pam.365, days.pam.max) %>% 
  reduce(left_join, by = c("site", "richness.max")) %>% 
  mutate(site = factor(site, levels = c("Rinyirru", "Undara", "Wambiana", "Mourachan",
                                        "Duval", "Tarcutta")))

sampling.table <- pam.rich.prop %>% 
  select(Site = site, "PAM (total richness)" = richness.max, "Days to max richness" = day,
         "PAM (proportion 28 days)" = prop.28,
         "PAM (proportion 90 days)" = prop.90,
         "PAM (proportion 180 days)" = prop.180,
         "PAM (proportion 365 days)" = prop.365) %>% 
  add_row(Site = "Average", "PAM (total richness)" = round(mean(pam.rich.prop$richness.max)),
          "Days to max richness" = round(mean(pam.rich.prop$day)),
          "PAM (proportion 28 days)" = round(mean(pam.rich.prop$prop.28), digits = 2),
          "PAM (proportion 90 days)" = round(mean(pam.rich.prop$prop.90), digits = 2),
          "PAM (proportion 180 days)" = round(mean(pam.rich.prop$prop.180), digits = 2),
          "PAM (proportion 365 days)" = round(mean(pam.rich.prop$prop.365), digits = 2))
```

## Seasonal sampling variation

This data set has seasons defined as data within calendar seasons across all years.

Dec-Mar = Summer
Mar-Jun = Autumn
Jun-Sep = Winter
Sep-Dec = Spring

```{r}
frog_data_season <- read_csv("frog_data_season.csv")
```

```{r}
frog_community_season <- frog_data_season %>% 
  select(site, SamplingDay, season,scientific.name) %>% #selected relevant grouping variables
  group_by(across(everything())) %>% #groups by all available columns
  mutate(number = n()) %>% #created a numbers column that counts the number of duplicate observations
  ungroup() %>% 
  unique() %>% #merged the duplicate rows
  pivot_wider(names_from = "scientific.name", #transformed dataset from long to wide format
              values_from = "number",
              values_fill = list(number=0)) %>% #replaced NAs with 0
  arrange(site) %>% 
  mutate(SamplingDay = as.factor(SamplingDay))

frog_community_season2 <- frog_community_season %>%
    group_by(site, SamplingDay) %>%
    summarise(
        season = "total",
        across(where(is.numeric), sum),
        .groups = "drop"
    ) %>%
    bind_rows(frog_community_season, .)

```

```{r}
# A. Define the Seasonal Limits
seasonal_limits <- tribble(
  ~site,       ~season,   ~max_days,
  "Duval",     "winter",  147,
  "Mourachan", "winter",  184,
  "Rinyirru",  "winter",  110,
  "Tarcutta",  "winter",  276,
  "Undara",    "winter",  250,
  "Wambiana",  "winter",  276,
  
  "Duval",     "spring",  168,
  "Mourachan", "spring",  214,
  "Rinyirru",  "spring",  119,
  "Tarcutta",  "spring",  271,
  "Undara",    "spring",  227,
  "Wambiana",  "spring",  277,
  
  "Duval",     "summer",  180,
  "Mourachan", "summer",  200,
  "Rinyirru",  "summer",  147,
  "Tarcutta",  "summer",  177,
  "Undara",    "summer",  122,
  "Wambiana",  "summer",  361,
  
  "Duval",     "autumn",  184,
  "Mourachan", "autumn",  183,
  "Rinyirru",  "autumn",  92,
  "Tarcutta",  "autumn",  256,
  "Undara",    "autumn",  134,
  "Wambiana",  "autumn",  320
)


# B. Calculate "Total" Limits (Sum of all seasons per site)
total_limits <- seasonal_limits %>%
  group_by(site) %>%
  summarise(max_days = sum(max_days)) %>%
  mutate(season = "total") # Match the 'total' label in your data

# C. Combine into one Master Lookup
all_day_limits <- bind_rows(seasonal_limits, total_limits)

```

```{r}
#| fig-width: 14
#| fig-asp: 0.6
#| warning: false

incidence_data <- frog_community_season2 %>%
  # A. JOIN THE LIMITS (Attaches the correct 'max_days' to every row)
  left_join(all_day_limits, by = c("site", "season")) %>%
  # Create ID
  unite("Group_ID", site, season, sep = "__", remove = FALSE) %>%
  group_by(Group_ID, site, season) %>%
  # Convert counts to binary Presence(1)/Absence(0)
  mutate(across(where(is.numeric) & !c(SamplingDay, max_days), ~ ifelse(. > 0, 1, 0))) %>%
  summarise(
    # B. Extract the Limit (it's already in the column from the join)
    T_days = first(max_days),
    # C. Sum Species Counts (Exclude metadata columns!)
    across(where(is.numeric) & !c(SamplingDay, max_days), sum),
    .groups = "drop"
  )

# Create List
list_data <- split(incidence_data, incidence_data$Group_ID)

# Vector Cleaning Function
clean_vector <- function(x) {
  # 1. Extract T explicitly
  t_val <- x$T_days
  
  # 2. Extract Species Counts (Remove T_days from selection)
  counts <- x %>% 
    select(where(is.numeric), -T_days) %>% 
    unlist() %>% 
    as.numeric()
  
  # 3. Return c(T, count1, count2...)
  # Filter out 0s (species not seen in that season)
  return(unname(c(t_val, counts[counts > 0])))
}

# Apply Function
iNEXT_input <- lapply(list_data, clean_vector)

# Run iNEXT
# This will now calculate curves based on 147 days, 180 days, etc. correctly
out <- iNEXT(iNEXT_input, q = 0, datatype = "incidence_freq", knots = 100, endpoint = 450)

# Extract Results for Plotting
season_results <- out$iNextEst$size_based %>%
  as_tibble() %>%
  filter(Order.q == 0) %>%
  rename(x = t, y = qD, y.lwr = qD.LCL, y.upr = qD.UCL) %>%
  separate(Assemblage, into = c("site", "season"), sep = "__") %>%
  mutate(
    season = str_to_title(season),
    site = factor(site, levels = c("Rinyirru","Undara","Wambiana","Mourachan","Duval","Tarcutta")),
    season = factor(season, levels = c("Spring", "Summer", "Autumn", "Winter", "Total")),
    Method = ifelse(Method == "Observed", "Rarefaction", Method),
    Method = factor(Method, levels = c("Rarefaction", "Extrapolation"))
  )

(rare.inext_season <- ggplot(season_results, aes(x = x, y = y, colour = season)) +
  geom_ribbon(aes(ymin = y.lwr, ymax = y.upr, fill = season),
              alpha = 0.15, colour = NA) +
  geom_line(aes(linetype = Method), linewidth = 1.3) +
  facet_wrap(~ site, scales = "free_y") +
  scale_x_continuous(name = "Survey Effort (Days)",
                       limits = c(0,450),
                       breaks = seq(0, 450, by = 90)) +
  scale_y_continuous(name = "Species Richness") +
  scale_colour_manual(values = c("goldenrod1", "#F38181","#943CE7", "burlywood3", "black"),
                        name = "Sampling Period",
                      labels = c("Spring", "Summer", "Autumn", "Winter", "All season")) +
  scale_fill_manual(values = c("goldenrod1", "#F38181","#943CE7", "burlywood3", "black"),
                        guide = "none") +
  scale_linetype_manual(values = c(Rarefaction = "solid", Extrapolation = "22"),
                        guide = guide_legend(title = NULL)) +
 my.theme() +
  theme(
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1.5),
    strip.text.x = element_text(size = 15, face = "bold", hjust = 0),
    strip.background = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(face = "italic", size = 14)
  ) +
  guides(colour = guide_legend(order = 1),
         linetype = guide_legend(order = 2)))

```


## Days to match OBM

Calculate how long it takes for each seasonal start to match OBM

```{r}
# 2. Define the Function
run_inext_seasonal <- function(group_id, data_vec) {
  
  # A. Look up the max days for this specific Site/Season combo
  max_d <- seasonal_limits[[group_id]]
  
  # B. Create the explicit sequence of days
  # LOGIC: Go up to 450, OR the actual max days if the season is longer than 450.
  # This ensures short seasons extrapolate to 450, but long seasons (Total) aren't cut short.
  target_end <- max(c(max_d, 450), na.rm = TRUE)
  
  day_seq <- seq(1, target_end, by = 1)
  
  # C. Run iNEXT
  res <- iNEXT(list(data_vec), q = 0, datatype = "incidence_freq", size = day_seq)
  
  # D. Extract dataframe
  df <- res$iNextEst$size_based %>%
    as_tibble() %>%
    filter(Order.q == 0) %>%
    mutate(Group_ID = group_id) # Attach ID to split later
    
  return(df)
}

# 3. Run the Loop
all_seasons_result <- map2_dfr(names(iNEXT_input), iNEXT_input, run_inext_seasonal) %>%
  
  # 4. Clean and Format
  rename(day = t, richness = qD, lower.ci = qD.LCL, upper.ci = qD.UCL) %>%
  
  # Split the ID back into Site and Season
  separate(Group_ID, into = c("site", "season"), sep = "__") %>%
  
  # Apply Factors for Plotting
  mutate(
    season = str_to_title(season),
    season = recode(season, "Total" = "All season"), # Rename Total here
    
    site = factor(site, levels = c("Rinyirru", "Undara", "Wambiana", "Mourachan",
                                   "Duval", "Tarcutta")),
    season = factor(season, levels = c("Spring", "Summer", "Autumn", "Winter", "All season")),
    Method = ifelse(Method == "Observed", "Rarefaction", Method)
  )

```


```{r}
days_to_reach_season <- all_seasons_result %>%
  left_join(obm.max, by = "site") %>%
  group_by(site,season) %>%
  filter(round(richness, digits = 1) >= richness.max) %>%
  slice(1) %>%
  select(site, season, days_required = day, pam_richness = richness, obm_target = richness.max)
```


# Session Info
```{r}
sessionInfo()
```

# Cite Packages
```{r}
cite_packages()
```
